---
title: "Practical Machine Learning Project"
author: "E.T.Jaba Jasphin"
date: "June 21, 2015"
output: html_document
---

In this project, my goal is to determine how well the participants are doing the barbell lifts and to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. 
The required Libraries are loded below

```{r, echo=TRUE,message=FALSE} 
library(caret)
library(ggplot2)
library(kernlab)
set.seed(33245)
```
For reading only the required data the following commands were used

```{r, echo=TRUE,message=FALSE}
training<-read.csv("pml-training.csv")
testing<-read.csv("pml-testing.csv")

trainingdata<-c("accel_belt_x","accel_belt_y","accel_belt_z","accel_arm_x","accel_arm_y","accel_arm_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","accel_forearm_x","accel_forearm_y","accel_forearm_z","user_name","classe")
testingdata<-c("accel_belt_x","accel_belt_y","accel_belt_z","accel_arm_x","accel_arm_y","accel_arm_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","accel_forearm_x","accel_forearm_y","accel_forearm_z","user_name")

training1<-training[,trainingdata]
testing1<-testing[,testingdata]
```

Since our goal is to determine how well a participant is doing the barbell lifts, I have separated the data based on the user name and Created data partition for training and testing. 
```{r, echo=TRUE,message=FALSE}
x3<-subset(training1,user_name=="charles")
trainingcharles1<-x3[,-13]
inTraincharles<-createDataPartition(y=trainingcharles1$classe,p=0.75,list=FALSE);

trainingcharles<-trainingcharles1[inTraincharles,]
testingcharles<-trainingcharles1[-inTraincharles,]
```
Feature plot was created for some parameters which i have chosen for Charles.
```{r, echo=TRUE,message=FALSE}
featurePlot(x=trainingcharles[,c("accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z","accel_forearm_x","accel_forearm_y")],y=trainingcharles$classe)
```

Model was created using the method rpart and gbm. If iam choosing rpart method accuracy was very less.So i have chosen the method gbm.For cross validation K-fold method was chosen and preprocessing was used with center and scale method.

```{r, echo=TRUE,message=FALSE}
fitControl<-trainControl(method="repeatedcv",number=10,repeats=10)

modelFitcharles<-train(classe~.,data=trainingcharles,preProcess=c("center","scale"),trControl=fitControl,method="gbm",verbose=FALSE);
outputcharles<-predict(modelFitcharles,newdata=testingcharles)
confusionMatrix(testingcharles$classe,outputcharles)
```

Output for the person Charles in the given testing set is predicted as follows
```{r, echo=TRUE,message=FALSE}
charles_classe<-predict(modelFitcharles,newdata=testing1[10,])
print(charles_classe)
```
The above is the predicted classe for Charles in the testing set.

